Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-01-25 WSSSPE3/report/wg_main_funding_spec_expert.tex
Fix typos and minor rewriting.

Research Software Engineers (RSEs) -- those who contribute to science and
scholarship through software development -- are an important part of the team
needed to deliver 21st century research. However, existing academic structures
and systems of funding do not effectively fund and sustain these skills. The
resulting high levels of turnover and inappropriate incentives are significant
contributing factors to low levels of reliability and readability observed in
scientific software. Moreover, the absence of skilled and experienced developers
retards progress in key projects, and at times causes important projects to fail
completely.

Effective development of software for advanced research requires that
researchers work closely with scientific software developers who understand the
research domain sufficiently to build meaningful software at a reasonable pace.
This requires a collaborative approach -- where developers who are fully engaged
or invested in the research context are co-developing software with domain
academics.

@@

One solution is to fund dedicated software engineering roles for major research
software projects at national laboratories or other non-educational
institutions. This solution is in place and working well for many well-used
scientific codebases. However, this strategy has limited application, as much of
the body of software is created and maintained in research universities.
Therefore, we argue that research institutions should develop hybrid
academic-technical tracks for this capacity, where employees in this track work
with more than one PI, rather than the traditional RA role within a single
group. This could be coordinated centrally, as a core facility, perhaps within
research computing organizations which have traditionally supported university
cyberinfrastructure, library organizations, or research offices. Alternatively,
these groups could be organizationally closer to research groups, sitting within
academic departments. The most effective model will vary from institution to
institution, but the mandate and ways of working should be similar.

>>>>

Research Software Engineers (RSEs)---those who contribute to science and
scholarship through software development---are an important part of the team
needed to deliver 21st century research. However, existing academic structures
and systems of funding do not effectively fund and sustain these skills. The
resulting high levels of turnover and inappropriate incentives are significant
contributing factors to low levels of reliability and readability observed in
scientific software. Moreover, the absence of skilled and experienced developers
retards progress in key projects, and at times causes important projects to fail
completely.

Effective development of software for advanced research requires that
researchers work closely with scientific software developers who understand the
research domain sufficiently to build meaningful software at a reasonable pace.
This requires a collaborative approach---where developers who are fully engaged
or invested in the research context are co-developing software with domain
academics.

@@

One solution is to fund dedicated software engineering roles for major research
software projects at national laboratories or other non-educational
institutions. This solution is in place and working well for many well used
scientific codebases. However, this strategy has limited application, as much of
the body of software is created and maintained in research universities.
Therefore, we argue that research institutions should develop hybrid
academic-technical tracks for this capacity, where employees in this track work
with more than one PI, rather than the traditional RA role within a single
group. This could be coordinated centrally, as a core facility, perhaps within
research computing organizations which have traditionally supported university
cyberinfrastructure, library organizations, or research offices. Alternatively,
these groups could be organizationally closer to research groups, sitting within
academic departments. The most effective model will vary from institution to
institution, but the mandate and ways of working should be similar.

### 2016-01-25 WSSSPE3/report/wg_main_metrics.tex
Fix typos and minor rewriting.

Metrics for scientific software are important for tenure and promotion,
scientific impact, discovery, reducing duplication, serving as a basis for
potential industrial interest in adopting software, prioritizing development and
support towards strategic objectives, and making a case for new or continued
funding. However, there is no commonly-used standard for collecting or
presenting metrics, nor is it known if there is a common set of metrics for
scientific software. It is imperative that scientific software stakeholders
understand that it is useful to collect metrics.

>>>>

Metrics for scientific software are important for tenure and promotion,
scientific impact, discovery, reducing duplication, serving as a basis for
potential industrial interest in adopting software, prioritizing development and
support towards strategic objectives, and making a case for new or continued
funding. However, there is no commonly used standard for collecting or
presenting metrics, nor is it known if there is a common set of metrics for
scientific software. It is imperative that scientific software stakeholders
understand that it is useful to collect metrics.

### 2016-01-25 WSSSPE3/report/wg_main_training.tex
Fix typos and minor rewriting.

This group explored a rapidly growing array of training that is seen to
contribute to sustainable software. The offerings are diverse, providing
training that is more or less directly relevant to sustainable software. While
research institutions support professional development for research staff, the
skills taught which might impact on sustainable software are limited at best,
often lacking a clear and coherent development pathway. This growing array of
training opportunities could usefully be coordinated by bringing together those
involved in leading relevant initiatives on a regular basis.

>>>>

This group explored a rapidly growing array of training that is seen to
contribute to sustainable software. The offerings are diverse, providing
training that is more or less directly relevant to sustainable software. While
research institutions support professional development for research staff, the
skills taught which might impact on sustainable software are limited at best,
often lacking a clear and coherent development pathway. Bringing together those
involved in leading relevant initiatives on a regular basis could helpfully
coordinate this growing array of training opportunities.
